A Mechanised Library for Generic (Co)Programming

### 2024-02-23 paper/paper.tex
a bit more explanation of structural vs nonstructural recursion in section 5

Termination and productivity are non-trivial properties of programs, and various
methods have been proposed for checking that these properties hold. A common
approach is guardedness checking~\cite{10.5555/646535.695850}, which is a
\emph{syntactic} check that definitions avoid the introduction of
non-normalizable terms. However, some desirable definitions will not be accepted
by such checks.

In particular, the problem of \emph{nonstructural recursion} is
well-studied~\cite{BOVE_KRAUSS_SOZEAU_2016}.
% A distinction between standard and
% nonstandard recursion lies in the fact that standard recursion adheres strictly
% to recursing on pattern variables, avoiding reliance on results from other
% function calls. Nonstandard recursion may deviate from this convention, allowing
% for more flexibility.
Certain functions that are not structurally recursive can be reformulated using
a nonstandard approach to achieve structural recursion~\cite{AbreuDHJMS23}.
Take, for instance, division by iterated subtraction, which is inherently
non-structurally recursive since it involves recursion based on the result of a
subtraction. There is a nonstandard implementation of divivion found on Coq's
standard library, which involves a four-argument function that effectively
combines subtraction and division. Similarly, the mergesort in Coq's standard
library uses an ``explicit stack of pending merges'' in order to avoid issues
with nonstructural definitions. There is a major downside, however; as noted by
Abreu et al., the result is ``barely recognizable as a form of
mergesort''~\cite{AbreuDHJMS23}.

>>>>

Termination and productivity are non-trivial properties of programs, and various
methods have been proposed for checking that these properties hold. A common
approach is guardedness checking~\cite{10.5555/646535.695850}, which is a
\emph{syntactic} check that definitions avoid the introduction of
non-normalizable terms. This sort of check generally looks for a
\emph{structural decrease} of arguments in the recursive calls of a function.
Coq uses such a check, and it works for many classic functional programming
patterns (like \haskell{map} and \haskell{foldr}). However, some desirable
definitions will not be accepted by such checks.

In particular, the problem of \emph{nonstructural recursion} (including
divide-and-conquer algorithms) is well-studied~\cite{BOVE_KRAUSS_SOZEAU_2016}.
% A distinction between standard and
% nonstandard recursion lies in the fact that standard recursion adheres strictly
% to recursing on pattern variables, avoiding reliance on results from other
% function calls. Nonstandard recursion may deviate from this convention, allowing
% for more flexibility.
Certain functions that are not structurally recursive can be reformulated using
a nonstandard approach to achieve structural recursion~\cite{AbreuDHJMS23}.
Take, for instance, division by iterated subtraction, which is inherently
non-structurally recursive since it involves recursion based on the result of a
subtraction. There is a nonstandard implementation of divivion found on Coq's
standard library, which involves a four-argument function that effectively
combines subtraction and division. Similarly, the mergesort in Coq's standard
library uses an ``explicit stack of pending merges'' in order to avoid issues
with nonstructural definitions. There is a major downside, however; as noted by
Abreu et al., the result is ``barely recognizable as a form of
mergesort''~\cite{AbreuDHJMS23}.
